---
title: "p8105_hw2_at3346"
author: "Ashley Tseng"
date: "10/4/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readxl)
library(formattable)
```

## Problem 1
Read and clean the Mr. Trash Wheel sheet:
```{r problem1.1}
trashwheel_data = 
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "Mr. Trash Wheel", col_names = TRUE, range = "A2:N406") %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = round(sports_balls, digits = 0),
    sports_balls = as.integer(sports_balls))
```

Read and clean precipitation data for 2017 and 2018:
```{r problem1.2}
precip_2017 = 
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "2017 Precipitation", col_names = TRUE, range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  drop_na(total) %>% 
  mutate(year = "2017")

precip_2018 = 
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "2018 Precipitation", col_names = TRUE, range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  drop_na(total) %>% 
  mutate(year = "2018")
```

Combine precipitation datasets for 2017 and 2018 & convert month to a character variable:
```{r problem1.3}
precip_both = 
  full_join(precip_2017, precip_2018, by = c("year", "month", "total")) %>% 
  mutate(month = month.name[month])
```
The number of observations in the Mr. Trash Wheel dataset is `r nrow(trashwheel_data)` observations. The average weight of all dumpsters is `r round(mean(trashwheel_data$weight_tons, digits = 2))` tons for available data. The average volume of trash is `r round(mean(trashwheel_data$volume_cubic_yards, digits = 2))` cubic yards for available data. The weight and volume of dumpsters is important because it reflects how much trash is being collected in Baltimore's Inner Harbor. The number of observations in the combined precipitation dataset is `r nrow(precip_both)` observations. For available data, the total precipitation in 2018 was `r round(sum(precip_2018$total), digits = 2)` inches. The median number of sports balls in a dumpster in 2017 was `r median(trashwheel_data$sports_balls)` balls.



## Problem 2
Clean the data in pols-month.csv:
```{r problem2.1}
pols_data = 
  read_csv("./data/pols-month.csv", col_names = TRUE) %>% 
  separate(mon, into = c("year","month", "day")) %>% 
  mutate(month = month.name[as.numeric(month)],
         president = recode(prez_gop, "1" = "gop", "2" = "gop", "0" = "dem")) %>% 
  transform(year = as.numeric(year)) %>% 
  select(-c("day", "prez_gop", "prez_dem"))
```

Clean the data in snp.csv:
```{r problem2.2}
snp_data = 
  read_csv("./data/snp.csv", col_names = TRUE) %>% 
  separate(date, into = c("month","day", "year")) %>% 
  mutate(month = month.name[as.numeric(month)]) %>% 
  transform(year = as.numeric(year)) %>% 
  select(year, month, everything(),-c("day"))
```

Clean the data in unemployment.csv:
```{r problem2.3}
unemployment_data = 
  read_csv("./data/unemployment.csv", col_names = TRUE) %>%
  pivot_longer(
    cols = Jan:Dec,
    names_to = "month",
    values_to = "unemployment_pct") %>% 
  mutate(month = match(month, month.abb),
         month = month.name[month]) %>% 
  janitor::clean_names()
```

Join snp_data to pols_data:
```{r problem2.4}
snp_pols_data = left_join(pols_data, snp_data, by = c("year", "month"))
```

Join unemployment_data to snp_pols_data:
```{r problem2.5}
un_snp_pols_data = left_join(snp_pols_data, unemployment_data, by = c("year", "month"))
```

`pols_data` contains `r nrow(pols_data)` observations and `r ncol(pols_data)` variables related to the number of national politicians (the president, governors, senators, and representatives) who are democratic or republican between `r min(pols_data$year)` and `r max(pols_data$year)`. `snp_data` contains `r nrow(snp_data)` observations and `r ncol(snp_data)` variables related to Standard & Poor’s stock market index (S&P) between `r min(snp_data$year)` and `r max(snp_data$year)`, often used as a representative measure of stock market as a whole. `unemployment_data` contains `r nrow(unemployment_data)` observations and `r ncol(unemployment_data)` variables, providing the percentage of unemployment in a particular month of the associated year between `r min(unemployment_data$year)` and `r max(unemployment_data$year)`. The aggregated dataset of `pols_data`, `snp_data`, and `unemployment_data` is named `un_snp_pols_data` which contains `r nrow(un_snp_pols_data)` observations and `r ncol(un_snp_pols_data)` variables between `r min(un_snp_pols_data$year)` and `r max(un_snp_pols_data$year)`. Key variables in `un_snp_pols_data` include "president," "closing," and "unemployment_pct". "president" is a key variable since it indicates the political party of the president which is interesting given the distribution of democratic and republican politicians in office at the time. "closing" is a key variable since it provides the closing values of the S&P stock index on the associated date of observation. "unemployment_pct" is a key variable because it shows the unemployment percentage.



## Problem 3

Load and clean the data in Popular_Baby_Names.csv:
```{r problem3.1}
pop_baby_data = 
  read_csv("./data/Popular_Baby_Names.csv", col_names = TRUE) %>% 
  janitor::clean_names() %>% 
  transform(
    gender = tolower(gender),
    ethnicity = tolower(ethnicity),
    childs_first_name = stringr::str_to_title(childs_first_name)) %>% 
  mutate(
    ethnicity = if_else(ethnicity == 'asian and paci', 'asian and pacific islander', ethnicity),
    ethnicity = if_else(ethnicity == 'black non hisp', 'black non hispanic', ethnicity),
    ethnicity = if_else(ethnicity == 'white non hisp', 'white non hispanic', ethnicity)) %>% 
  distinct()
```

Produce a table showing the rank in popularity of the name “Olivia” as a female baby name over time:
```{r problem3.2}
olivia_name = 
  pop_baby_data %>% 
  filter(gender == "female", childs_first_name == "Olivia") %>% 
  group_by(ethnicity, year_of_birth) %>% 
  summarize(rank) %>% 
  pivot_wider(
    names_from = year_of_birth,
    values_from = rank
  ) %>% 
  knitr::kable()

olivia_name
```

Produce a table showing the most popular name among male children over time:
```{r problem3.3}
male_children = 
  pop_baby_data %>% 
  filter(gender == "male", rank == 1) %>% 
  group_by(ethnicity, year_of_birth) %>% 
  summarize(childs_first_name) %>% 
  pivot_wider(
    names_from = year_of_birth,
    values_from = childs_first_name
  )%>% 
  knitr::kable()

male_children
```

For male, white non-hispanic children born in 2016, produce a scatter plot showing the number of children with a name (y axis) against the rank in popularity of that name (x axis):
```{r problem3.4}
male_white_2016 = 
  pop_baby_data %>% 
  filter(gender == "male", ethnicity == "white non hispanic", year_of_birth == 2016)

mw2016_plot = 
  male_white_2016 %>% 
  ggplot(aes(x = rank, y = count)) + 
  geom_point() +
  labs(
    title = "Popularity of Names of Male, White Non-hispanic Children Born in 2016",
    x = "Rank in Popularity of Name",
    y = "Number of Children with Name")

mw2016_plot

ggsave("mw2016_plot.pdf")

```

